---
title: "Music Perception 1: Variability Analysis"
output: html_notebook
---

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

- 21 participants (experimental error = 1) >> 20 participants were included for analysis.

# Data analysis
- Pearson's correlation (if normality is violated, Spearman's rank correlation instead) between variability measurements and participants' judgments as "teaching"

## Variables
- CV for IOIs (tempo)
- SD for KOT (articulation)
- CV for KOT (articulation)
- CV for KV (dynamics)

```{r packages, include = FALSE}
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# midi
if (!require("tuneR")) {install.packages("tuneR"); require("tuneR")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
```

```{r read, include = FALSE}
# read files and combine them
data_ls <- list.files("data", pattern = "csv")
combined <- lapply(data_ls, function(f){
  fread(paste("data/", f, sep = ""), sep = ",")
})
all_data <- do.call(rbind.data.frame, combined)

# remove irrelevant strings
data <- all_data[expMode == "experiment"]
data$MidFile <- gsub("./mid/", "", data$midFile)
data$MidFile <- gsub(".mid", "", data$MidFile)

# name conditions
data$Condition <- "NA"
data$Skill <- "NA"
data[grepl("_t_", midFile)]$Condition <- "teaching"
data[grepl("_p_", midFile)]$Condition <- "performing"
data[grepl("a_", midFile)]$Skill <- "articulation"
data[grepl("d_", midFile)]$Skill <- "dynamics"
```

# Judged as "teaching" (%) for each stimulus
```{r teaching, echo = FALSE}
data$Teaching <- 0
data[rating == "Yes"]$Teaching <- 1

# individual
individual <- data[subjectNumber != 9, .(N = .N, Sum = sum(Teaching)), by = .(subjectNumber, Condition, Skill, MidFile)]
individual<- individual[order(subjectNumber)]

# all
all <- individual[, .(N = .N, Mean = mean(Sum), SD = sd(Sum), Mean_Percent = (mean(Sum))*100, SD_Percent = (sd(Sum))*100), by = .(Condition, Skill, MidFile)]
all <- all[order(MidFile)]
all
```

```{r midi, include = FALSE}
# create a list of data file names
lf <- list.files("../experiment/mid/", pattern = "mid")

# combine data
all_data <- data.table()
for (file in lf){
   current <- readMidi(paste("../experiment/mid/", file, sep = ""))
   current$MidFile <- gsub(".mid", "", file)
   all_data <- rbind(all_data, current)
}

data <- all_data[event == "Note On" | event == "Note Off"]
data$NoteNr <- rep(1:134, nrow(data)/134)
data$channel <- NULL
data$parameterMetaSystem <- NULL
data$track <- NULL
data$type <- NULL
data$time2 <- data$time
#data$time2 <- (data$time/960)*1000 # deltatime to time (ms)

# labelling
colnames(data)[c(3,4)] <- c("Pitch", "Velocity") # time event also change
data$Condition <- "NA"
data$Skill <- "NA"
data[grepl("_t_", MidFile)]$Condition <- "teaching"
data[grepl("_p_", MidFile)]$Condition <- "performing"
data[grepl("a_", MidFile)]$Skill <- "articulation"
data[grepl("d_", MidFile)]$Skill <- "dynamics"

# define subcomponents
# For intervals
ls_legato <- list(c(1:7), c(17:23), c(42:48), c(58:64))
ls_staccato <- list(c(9:15), c(25:31), c(34:40), c(50:56))
ls_forte <- list(c(1:7), c(17:23), c(42:48), c(58:64))
ls_piano <- list(c(9:15), c(25:31), c(34:40), c(50:56))

# For each note (velocity only)
ls_legato_2 <- list(c(1:8), c(17:24), c(42:49), c(58:65))
ls_staccato_2 <- list(c(9:16), c(25:32), c(34:41), c(50:57))
ls_forte_2 <- list(c(1:8), c(17:24), c(42:49), c(58:65))
ls_piano_2 <- list(c(9:16), c(25:32), c(34:41), c(50:57))

# Define Skill Change (LtoS, FtoP)
change_1 <- c(8, 24, 49)
# Define Skill Change (StoL, PtoF)
change_2 <- c(16, 41, 57)
```

# Results
- significance levels: `***: 0.001, **: 0.01, *: 0.05`
- scatter plots with linear regression lines and 95% confidence intervals (gray)
- SD: Standard deviation
- CV: Coefficient of variation (SD/mean)

# IOIs: CV (tempo variability)
```{r ioi, include = FALSE}
data_onset <- data[event == "Note On"]
data_offset <- data[event == "Note Off"]

data_onset$IOI <- diff(c(0, data_onset$time2))
dt_ioi <- data_onset[NoteNr != 1]
dt_ioi$Interval <- rep(1:66, nrow(dt_ioi)/66)

# assign Subcomponents
dt_ioi$Subcomponent <- "NA"
# Legato
for (phrase in 1:length(ls_legato)){
 for (note in 1:length(ls_legato[[phrase]])){
   dt_ioi[Skill == "articulation" & Interval == ls_legato[[phrase]][note]]$Subcomponent <- "Legato"
 }
}
# Staccato
for (phrase in 1:length(ls_staccato)){
 for (note in 1:length(ls_staccato[[phrase]])){
   dt_ioi[Skill == "articulation" & Interval == ls_staccato[[phrase]][note]]$Subcomponent <- "Staccato"
 }
}

# Forte
for (phrase in 1:length(ls_forte)){
 for (note in 1:length(ls_forte[[phrase]])){
   dt_ioi[Skill == "dynamics" & Interval == ls_forte[[phrase]][note]]$Subcomponent <- "Forte"
 }
}
# Piano
for (phrase in 1:length(ls_piano)){
 for (note in 1:length(ls_piano[[phrase]])){
   dt_ioi[Skill == "dynamics" & Interval == ls_piano[[phrase]][note]]$Subcomponent <- "Piano"
 }
}

# assign Subcomponent Change
for (number in change_1){
 dt_ioi[Skill == "articulation" & Interval == number]$Subcomponent <- "LtoS"
 dt_ioi[Skill == "dynamics" & Interval == number]$Subcomponent <- "FtoP"
}
for (number in change_2){
 dt_ioi[Skill == "articulation" & Interval == number]$Subcomponent <- "StoL"
 dt_ioi[Skill == "dynamics" & Interval == number]$Subcomponent <- "PtoF"
}
```

## Whole excerpt
```{r cv-ioi, echo = FALSE}
rating <- all # rating data

cv <- dt_ioi[Subcomponent != "NA", .(N = .N, CV = sd(IOI)/mean(IOI)), by = .(MidFile, Skill)]
cv

# sorted
cv <- cv[order(MidFile)]

# plot
cv$Teaching <- rating$Mean_Percent
cv$TeachingSD <- rating$SD_Percent

ggscatter(cv, x = "CV", y = "Teaching", color = "Skill", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE,
          xlab = "CV", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

### Articulation (n.s.)
```{r, echo = FALSE}
# normality check
shapiro.test(cv[Skill == "articulation"]$Teaching)
shapiro.test(cv[Skill == "articulation"]$CV)

cor.test(cv[Skill == "articulation"]$Teaching, cv[Skill == "articulation"]$CV)

cor.test(cv[Skill == "articulation"]$Teaching, cv[Skill == "articulation"]$CV, method = "spearman", exact = FALSE)
```

### Dynamics (n.s.)
```{r, echo = FALSE}
# normality check
shapiro.test(cv[Skill == "dynamics"]$Teaching)
shapiro.test(cv[Skill == "dynamics"]$CV)

cor.test(cv[Skill == "dynamics"]$Teaching, cv[Skill == "dynamics"]$CV)

cor.test(cv[Skill == "dynamics"]$Teaching, cv[Skill == "dynamics"]$CV, method = "spearman", exact = FALSE)
```

## Except transition points
Generally pianists were likely to play slowly around transition points where one subcomponent of a technique moves to the other (e.g., from legat to staccato, from forte to piano). Therefore, we excluded these points to look at tempo variability.

```{r cv-ioi-2, echo = FALSE}
cv_2 <- dt_ioi[Subcomponent == "Legato" | Subcomponent == "Staccato" | Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, CV = sd(IOI)/mean(IOI)), by = .(MidFile, Skill)]
cv_2

# sorted
cv_2 <- cv_2[order(MidFile)]

# plot
cv_2$Teaching <- rating$Mean_Percent
cv_2$TeachingSD <- rating$SD_Percent

ggscatter(cv_2, x = "CV", y = "Teaching", color = "Skill", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE,
          xlab = "CV", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

### Articulation (n.s.)
```{r, echo = FALSE}
# normality check
shapiro.test(cv_2[Skill == "articulation"]$Teaching)
shapiro.test(cv_2[Skill == "articulation"]$CV)

cor.test(cv_2[Skill == "articulation"]$Teaching, cv_2[Skill == "articulation"]$CV)

cor.test(cv_2[Skill == "articulation"]$Teaching, cv_2[Skill == "articulation"]$CV, method = "spearman", exact = FALSE)
```

### Dynamics (*)
```{r, echo = FALSE}
# normality check
shapiro.test(cv_2[Skill == "dynamics"]$Teaching)
shapiro.test(cv_2[Skill == "dynamics"]$CV)

cor.test(cv_2[Skill == "dynamics"]$Teaching, cv_2[Skill == "dynamics"]$CV)

cor.test(cv_2[Skill == "dynamics"]$Teaching, cv_2[Skill == "dynamics"]$CV, method = "spearman", exact = FALSE)
```
# KOT: SD (articulation variability)
Because some recordings have a negative mean of KOT even for legato parts, SD, instead of CV, would be suitble to see articulation variation.

```{r kot, include = FALSE}
data_onset$KOT <- 0
for (row in 1:nrow(data_onset)){
   if (row < nrow(data_onset)){
      data_onset$KOT[row+1] <- data_offset$time2[row] - data_onset$time2[row+1] # offset(n) - onset(n+1)
   }
}

dt_kot <- data_onset[NoteNr != 1]
dt_kot$Interval <- rep(1:66, nrow(dt_kot)/66)

# assign Subcomponents
dt_kot$Subcomponent <- "NA"
# Legato
for (phrase in 1:length(ls_legato)){
   for (note in 1:length(ls_legato[[phrase]])){
     dt_kot[Skill == "articulation" & Interval == ls_legato[[phrase]][note]]$Subcomponent <- "Legato"
   }
}
# Staccato
for (phrase in 1:length(ls_staccato)){
   for (note in 1:length(ls_staccato[[phrase]])){
     dt_kot[Skill == "articulation" & Interval == ls_staccato[[phrase]][note]]$Subcomponent <- "Staccato"
   }
}
# Forte
for (phrase in 1:length(ls_forte)){
   for (note in 1:length(ls_forte[[phrase]])){
     dt_kot[Skill == "dynamics" & Interval == ls_forte[[phrase]][note]]$Subcomponent <- "Forte"
   }
}
# Piano
for (phrase in 1:length(ls_piano)){
   for (note in 1:length(ls_piano[[phrase]])){
     dt_kot[Skill == "dynamics" & Interval == ls_piano[[phrase]][note]]$Subcomponent <- "Piano"
   }
}
# assign Subcomponent Change
for (number in change_1){
   dt_kot[Skill == "articulation" & Interval == number]$Subcomponent <- "LtoS"
   dt_kot[Skill == "dynamics" & Interval == number]$Subcomponent <- "FtoP"
}
for (number in change_2){
   dt_kot[Skill == "articulation" & Interval == number]$Subcomponent <- "StoL"
   dt_kot[Skill == "dynamics" & Interval == number]$Subcomponent <- "PtoF"
}
```

## Articulation
```{r sd-kot-art, echo = FALSE}
sd_kot_art <- dt_kot[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(MidFile, Subcomponent)]
sd_kot_art

# sorted
sd_kot_art <- sd_kot_art[order(MidFile)]

# plot
sd_kot_art$Teaching <- rep(rating[Skill == "articulation"]$Mean_Percent, each = 2)
sd_kot_art$TeachingSD <- rep(rating[Skill == "articulation"]$SD_Percent, each = 2)

ggscatter(sd_kot_art, x = "SD", y = "Teaching", color = "Subcomponent", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE,
          xlab = "SD (ms)", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

### Legato (n.s.)
```{r, echo = FALSE}
# normality check
shapiro.test(sd_kot_art[Subcomponent == "Legato"]$Teaching)
shapiro.test(sd_kot_art[Subcomponent == "Legato"]$SD)

cor.test(sd_kot_art[Subcomponent == "Legato"]$Teaching, sd_kot_art[Subcomponent == "Legato"]$SD)
```

### Staccato (n.s.)
```{r, echo = FALSE}
# normality check
shapiro.test(sd_kot_art[Subcomponent == "Staccato"]$Teaching)
shapiro.test(sd_kot_art[Subcomponent == "Staccato"]$SD)

cor.test(sd_kot_art[Subcomponent == "Staccato"]$Teaching, sd_kot_art[Subcomponent == "Staccato"]$SD)

cor.test(sd_kot_art[Subcomponent == "Staccato"]$Teaching, sd_kot_art[Subcomponent == "Staccato"]$SD, method = "spearman", exact = FALSE)
```

## Dynamics
```{r sd-kot-dyn, echo = FALSE}
sd_kot_dyn <- dt_kot[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(MidFile, Subcomponent)]
sd_kot_dyn

# sorted
sd_kot_dyn <- sd_kot_dyn[order(MidFile)]

# plot
sd_kot_dyn$Teaching <- rep(rating[Skill == "dynamics"]$Mean_Percent, each = 2)
sd_kot_dyn$TeachingSD <- rep(rating[Skill == "dynamics"]$SD_Percent, each = 2)

ggscatter(sd_kot_dyn, x = "SD", y = "Teaching", color = "Subcomponent", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE,
          xlab = "SD (ms)", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

### Forte (n.s.)
```{r, echo = FALSE}
# normality check
shapiro.test(sd_kot_dyn[Subcomponent == "Forte"]$Teaching)
shapiro.test(sd_kot_dyn[Subcomponent == "Forte"]$SD)

cor.test(sd_kot_dyn[Subcomponent == "Forte"]$Teaching, sd_kot_dyn[Subcomponent == "Forte"]$SD)

cor.test(sd_kot_dyn[Subcomponent == "Forte"]$Teaching, sd_kot_dyn[Subcomponent == "Forte"]$SD, method = "spearman", exact = FALSE)
```

### Piano (n.s.)
```{r, echo = FALSE}
# normality check
shapiro.test(sd_kot_dyn[Subcomponent == "Piano"]$Teaching)
shapiro.test(sd_kot_dyn[Subcomponent == "Piano"]$SD)

cor.test(sd_kot_dyn[Subcomponent == "Piano"]$Teaching, sd_kot_dyn[Subcomponent == "Piano"]$SD)

cor.test(sd_kot_dyn[Subcomponent == "Piano"]$Teaching, sd_kot_dyn[Subcomponent == "Piano"]$SD, method = "spearman", exact = FALSE)
```

# KOT: CV (articulation variability)
Because some recordings have the negative mean value or KOT even for Legato parts (i.e., relatively short legato), this measurement might not be useful.

## Articulation
### Legato
```{r cv-kot-leg, echo = FALSE}
cv_kot_leg <- dt_kot[Subcomponent == "Legato", .(N = .N, CV = sd(KOT)/mean(KOT)), by = .(MidFile)]
cv_kot_leg

# sorted
cv_kot_leg <- cv_kot_leg[order(MidFile)]

# plot
cv_kot_leg$Teaching <- rep(rating[Skill == "articulation"]$Mean_Percent, each = 1)
cv_kot_leg$TeachingSD <- rep(rating[Skill == "articulation"]$SD_Percent, each = 1)

ggscatter(cv_kot_leg, x = "CV", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE,
          xlab = "CV", ylab = "Judged as teaching (%)") + ylim(0, 100)
```


### Staccato
```{r cv-kot-sta, echo = FALSE}
cv_kot_sta <- dt_kot[Subcomponent == "Staccato", .(N = .N, CV = sd(KOT)/mean(KOT)), by = .(MidFile)]
cv_kot_sta

# sorted
cv_kot_sta <- cv_kot_sta[order(MidFile)]

# plot
cv_kot_sta$Teaching <- rep(rating[Skill == "articulation"]$Mean_Percent, each = 1)
cv_kot_sta$TeachingSD <- rep(rating[Skill == "articulation"]$SD_Percent, each = 1)

ggscatter(cv_kot_sta, x = "CV", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE,
          xlab = "CV", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

## Dynamics
### Forte
```{r cv-kot-for, echo = FALSE}
cv_kot_for <- dt_kot[Subcomponent == "Forte", .(N = .N, CV = sd(KOT)/mean(KOT)), by = .(MidFile)]
cv_kot_for

# sorted
cv_kot_for <- cv_kot_for[order(MidFile)]

# plot
cv_kot_for$Teaching <- rep(rating[Skill == "articulation"]$Mean_Percent, each = 1)
cv_kot_for$TeachingSD <- rep(rating[Skill == "articulation"]$SD_Percent, each = 1)

ggscatter(cv_kot_for, x = "CV", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE,
          xlab = "CV", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

### Piano
```{r cv-kot-pia, echo = FALSE}
cv_kot_pia <- dt_kot[Subcomponent == "Piano", .(N = .N, CV = sd(KOT)/mean(KOT)), by = .(MidFile)]
cv_kot_pia

# sorted
cv_kot_pia <- cv_kot_pia[order(MidFile)]

# plot
cv_kot_pia$Teaching <- rep(rating[Skill == "articulation"]$Mean_Percent, each = 1)
cv_kot_pia$TeachingSD <- rep(rating[Skill == "articulation"]$SD_Percent, each = 1)

ggscatter(cv_kot_pia, x = "CV", y = "Teaching", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE,
          xlab = "CV", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

# KV: CV (dynamics variability)
```{r vel, include = FALSE}
dt_vel <- data_onset
dt_vel$Note <- rep(1:67, nrow(dt_vel)/67)
# assign Subcomponents
# for each note
dt_vel$Subcomponent <- "NA"
# Legato
for (phrase in 1:length(ls_legato_2)){
   for (note in 1:length(ls_legato_2[[phrase]])){
     dt_vel[Skill == "articulation" & Note == ls_legato_2[[phrase]][note]]$Subcomponent <- "Legato"
   }
}
# Staccato
for (phrase in 1:length(ls_staccato_2)){
   for (note in 1:length(ls_staccato_2[[phrase]])){
     dt_vel[Skill == "articulation" & Note == ls_staccato_2[[phrase]][note]]$Subcomponent <- "Staccato"
   }
}
# Forte
for (phrase in 1:length(ls_forte_2)){
   for (note in 1:length(ls_forte_2[[phrase]])){
     dt_vel[Skill == "dynamics" & Note == ls_forte_2[[phrase]][note]]$Subcomponent <- "Forte"
   }
}
# Piano
for (phrase in 1:length(ls_piano_2)){
   for (note in 1:length(ls_piano_2[[phrase]])){
     dt_vel[Skill == "dynamics" & Note == ls_piano_2[[phrase]][note]]$Subcomponent <- "Piano"
   }
}
```

## Dynamics
```{r cv-vel-dyn, echo = FALSE}
cv_vel_dyn <- dt_vel[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, CV = sd(Velocity)/mean(Velocity)), by = .(MidFile, Subcomponent)]
cv_vel_dyn

# sorted
cv_vel_dyn <- cv_vel_dyn[order(MidFile)]

# plot
cv_vel_dyn$Teaching <- rep(rating[Skill == "dynamics"]$Mean_Percent, each = 2)
cv_vel_dyn$TeachingSD <- rep(rating[Skill == "dynamics"]$SD_Percent, each = 2)

ggscatter(cv_vel_dyn, x = "CV", y = "Teaching", color = "Subcomponent", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE,
          xlab = "CV", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

### Forte (n.s.)
```{r, echo = FALSE}
# normality check
shapiro.test(cv_vel_dyn[Subcomponent == "Forte"]$Teaching)
shapiro.test(cv_vel_dyn[Subcomponent == "Forte"]$CV)

cor.test(cv_vel_dyn[Subcomponent == "Forte"]$Teaching, cv_vel_dyn[Subcomponent == "Forte"]$CV)
```

### Piano (n.s.)
```{r, echo = FALSE}
# normality check
shapiro.test(cv_vel_dyn[Subcomponent == "Piano"]$Teaching)
shapiro.test(cv_vel_dyn[Subcomponent == "Piano"]$CV)

cor.test(cv_vel_dyn[Subcomponent == "Piano"]$Teaching, cv_vel_dyn[Subcomponent == "Piano"]$CV)

cor.test(cv_vel_dyn[Subcomponent == "Piano"]$Teaching, cv_vel_dyn[Subcomponent == "Piano"]$CV, method = "spearman", exact = FALSE)
```

## Articulation
```{r cv-vel-art, echo = FALSE}
cv_vel_art <- dt_vel[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, CV = sd(Velocity)/mean(Velocity)), by = .(MidFile, Subcomponent)]
cv_vel_art

# sorted
cv_vel_art <- cv_vel_art[order(MidFile)]

# plot
cv_vel_art$Teaching <- rep(rating[Skill == "articulation"]$Mean_Percent, each = 2)
cv_vel_art$TeachingSD <- rep(rating[Skill == "articulation"]$SD_Percent, each = 2)

ggscatter(cv_vel_art, x = "CV", y = "Teaching", color = "Subcomponent", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE,
          xlab = "CV", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

### Legato (**)
```{r, echo = FALSE}
# normality check
shapiro.test(cv_vel_art[Subcomponent == "Legato"]$Teaching)
shapiro.test(cv_vel_art[Subcomponent == "Legato"]$CV)

cor.test(cv_vel_art[Subcomponent == "Legato"]$Teaching, cv_vel_art[Subcomponent == "Legato"]$CV)

cor.test(cv_vel_art[Subcomponent == "Legato"]$Teaching, cv_vel_art[Subcomponent == "Legato"]$CV, method = "spearman", exact = FALSE)
```

### Staccato (*)
```{r, echo = FALSE}
# normality check
shapiro.test(cv_vel_art[Subcomponent == "Staccato"]$Teaching)
shapiro.test(cv_vel_art[Subcomponent == "Staccato"]$CV)

cor.test(cv_vel_art[Subcomponent == "Staccato"]$Teaching, cv_vel_art[Subcomponent == "Staccato"]$CV)
```
